### 2024-01-16 theart.qmd
quarto doc template

---
title: "Preoperative Cognitive Profile Predictive of Cognitive Decline after Subthalamic Deep Brain Stimulation in Parkinson’s Disease"
shorttitle: "Cognition in PD after STN DBS"
author:
  - name: Josef Mana
    corresponding: false
    orcid: "0000-0002-7817-3978"
    email: "josef.mana@protonmail.com"
    role:
      - Conceptualization
      - Data curation
      - Investigation
      - Formal analysis
      - Methodology
      - Software
      - Visualization
      - Writing - original draft
      - Writing - review & editing
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Ondrej Bezdicek
    corresponding: true
    orcid: "0000-0002-5108-0181"
    email: "ondrej.bezdicek@gmail.com"
    role:
      - Conceptualization
      - Data curation
      - Investigation
      - Methodology
      - Supervision
      - Writing - original draft
      - Writing - review & editing
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Filip Ruzicka
    corresponding: false
    email: "filr@seznam.cz"
    role:
      - Investigation
      - Writing - review & editing
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Anna Fecikova
    corresponding: false
    email: "Anna.Fecikova@vfn.cz"
    role:
      - Investigation
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Olga Klempirova
    corresponding: false
    email: "novakovaol@seznam.cz"
    role:
      - Investigation
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Tomas Nikolai
    corresponding: false
    email: "nikolai@centrum.cz"
    role:
      - Investigation
      - Writing - review & editing
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Tereza Uhrova
    corresponding: false
    email: "tereza.uhrova@vfn.cz"
    role:
      - Investigation
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Evzen Ruzicka
    corresponding: false
    email: "eruzi@lf1.cuni.cz"
    role:
      - Conceptualization
      - Funding acquisition
      - Investigation
      - Writing - review & editing
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Dusan Urgosik
    corresponding: false
    email: "urgosik@gmail.com"
    role:
      - Investigation
    affiliations:
      - id: "NHH"
        name: "Na Homolce Hospital, Prague, Czech Republic"
        department: "Department of stereotactic and radiation neurosurgery"
  - name: Robert Jech
    corresponding: false
    orcid: "0000-0002-9732-8947"
    email: "jech@cesnet.cz"
    role:
      - Conceptualization
      - Data curation
      - Funding acquisition
      - Investigation
      - Resources
      - Supervision
      - Writing - review & editing
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
abstract: ""
keywords:
  - Parkinson’s disease
  - dementia
format:
 apaquarto-pdf:
   documentmode: man
   a4paper: true
   pdf-engine: lualatex
   floatsintext: false
   keep-tex: false
---

{{< include _extensions/wjschne/apaquarto/_apa_title.qmd >}}

# Introduction

# Methods

# Results

# Discussion

# References