### 2024-06-12 theart.qmd
Uploaded article, introduction added

---
lang: en
title: "Updated Criteria for the Diagnostic Procedure for Parkinson’s Disease Dementia on Level I and their Validity in Deep Brain Stimulation Cohort"
shorttitle: "Parkinson's Disease Dementia Level I Criteria"
author:
  - name: Martina Kvapilova
    corresponding: false
    orcid: "0009-0007-4665-3946"
    role:
      - Conceptualization
      - Data curation
      - Writing - original draft
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Josef Mana
    corresponding: false
    orcid: "0000-0002-7817-3978" 
    email: "josef.mana@protonmail.com"
    role:
      - Conceptualization
      - Data curation
      - Investigation
      - Formal analysis
      - Software
      - Methodology
      - Project administration
      - Validation
      - Writing - original draft
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Martin Cihak
    corresponding: false
    orcid: "0000-0001-7765-4799"
    email: "martin.cih@gmail.com"
    role:
      - Conceptualization
      - Investigation
      - Methodology
      - Writing - original draft
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Tereza Uhrova
    corresponding: false
    email: "tereza.uhrova@vfn.cz"
    role:
      - Investigation
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Robert Jech
    corresponding: false
    orcid: "0000-0002-9732-8947"
    email: "jech@cesnet.cz"
    role:
      - Funding acquisition
      - Resources
      - Writing - review & editing
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Ondrej Bezdicek
    corresponding: true
    orcid: "0000-0002-5108-0181"
    email: "ondrej.bezdicek@gmail.com"
    role:
      - Investigation
      - Data curation
      - Funding acquisition
      - Conceptualization
      - Project administration
      - Supervision
      - Writing - original draft
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
abstract: ""
keywords:
  - Parkinson’s disease
  - Parkinson's disease dementia
  - level I criteria
format:
 apaquarto-docx:
   documentmode: man
   a4paper: true
   pdf-engine: lualatex
   floatsintext: false
   keep-tex: false
bibliography: references.bib
---

{{< include _extensions/wjschne/apaquarto/_apa_title.qmd >}}

# Introduction

Parkinson’s disease (PD) is a neurodegenerative disorder characterised by the slow, progressive onset of rigidity, bradykinesia, and resting tremors that ultimately advance to motor and non-motor impairment [@postuma2015].

The diagnostic criteria for Parkinson’s disease dementia (PDD) were developed in 2006 and their procedure followed shortly after [@emre2007; @dubois2007]. At that period the PDD criteria were oriented at criteria originating in Alzheimer’s disease (AD) since there was a lack of PD biomarkers to promote a biological classification of PD [@yamashita2023; @höglinger2024].

The newer Parkinson’s disease mild cognitive criteria (PD-MCI) led to an advance in different aspects of PD research [@litvan2012]. These included progress in clinical, genetic, therapeutic modalities, clinical trials or conversion rates to PDD [@hoogland2017; @hoogland2019; @aarsland2021].

At present, there is an ongoing effort to revisit the PDD criteria to secure the comparability of studies and clinical trials across different sites [@kulisevsky2024]. In our previous research, we aimed at establishing rigorous psychometric procedure to differentiate PD-MCI from PDD [@bezdicek2017; @bezdicek2016]. Hence the diagnosis of PDD is an exclusion criterion in deep brain stimulation (DBS), we decided to test the revisited PDD criteria in our PD cohort selected for undergoing DBS [@deuschl2006].

The current study aims at showing the agreement and differences between establishing the diagnosis using only Level 1 criteria as suggested by previous diagnostic procedure (the Movement Disorders Society (MDS)-Task Force screening checklist for detecting PD) [@emre2007; @dubois2007] in comparison to newly revised [@kulisevsky2024] one to validate them. The second objective of our study was to determine the agreement of revised PDD criteria with full neuropsychological testing at Level 2.

# Methods

## Participants

The data of patients with idiopathic PD diagnosed by a movement disorder specialist fulfilling the Movement Disorder Society (MDS) Clinical Diagnostic Criteria for Parkinson's disease (PD) [@postuma2015] were retrospectively gathered from clinical records acquired between January 2014 and December 2023. All the patients underwent a neuropsychological assessment by a trained clinical psychologist (OB) during a routine examination of cognitive functions as a part of the evaluation process for the indication of Deep Brain Stimulation (DBS) at General University Hospital in Prague. The Ethics Committee of the General University Hospital in Prague had approved the study protocol. All patients provided written informed consent prior to the examination.

## Neuropsychological Assessment

Participants were assessed with both MMSE [@stepankova2014; @folstein1975] and MoCA [@kopecek2016; @nasreddine2005] to measure overall cognitive performance. Moreover, a comprehensive neuropsychological assessment was performed in accordance with MDS Task Force Level 2 criteria for MCI in PD [@litvan2012]. We described our battery including a regression based calculator for normative scores in another study [@bezdicek2017]. Besides other neuropsychological tests, the comprehensive assessment employed Clock Drawing Test (CDT) and Letter Fluency. To establish cognitive performance in individual cognitive domains according to Level 1 criteria for PDD, we used test scores proposed by [@dubois2007] and their analogues in MoCA, see Table A for details.

```{r}
#| label: tbl-martin
#| tbl-cap: Comparison of criteia.
#| apa-note: The note below the table.

tibble::tribble(
  ~"Criteria", ~"Former", ~"New",
  "", "MMSE < 26", "MoCA < 26", 
  "Attention", "Serial 7's/months reversed", "Serial 7's",
  "Executive functions", "Lexical Fluency (S)/ Clock Drawing Test", "Lexical Fluency (K)/ Clock Drawing Test",
  "Visuo-spatial functions", "Drawing of Pentagons", "Drawing of Cube",
  "Memory", "3-Word Recall", "5-Word Recall",
  "Language", "-", "Naming of Animals",
  ) |>
  flextable::flextable() |>
  flextable::theme_apa()
```

To measure independence in activities of daily living, we administered the Functional Activities Questionnaire (FAQ) [@bezdicek2011; @pfeffer1982]. To assess neuropsychiatric functioning, we used Beck Depression Inventory-II (BDI) [@ciharova2020; @beck1996] and State-Trait Anxiety Inventory (STAI) [@spielberg1983; @mullner1980]. Psychotic symptoms were assessed in an interview by a trained psychiatrist.

# Results

# Discussion

# References