### 2025-01-05 texts/theart.qmd
Text update

---
lang: en
title: "Updated Criteria for the Diagnostic Procedure for Parkinson’s Disease Dementia on Level I and their Validity in Deep Brain Stimulation Cohort"
shorttitle: "Parkinson's Disease Dementia Level I Criteria"
author:
  - name: Martina Kvapilova
    corresponding: false
    orcid: "0009-0007-4665-3946"
    role:
      - Conceptualization
      - Data curation
      - Writing - original draft
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Josef Mana
    corresponding: false
    orcid: "0000-0002-7817-3978" 
    email: "josef.mana@protonmail.com"
    role:
      - Conceptualization
      - Data curation
      - Investigation
      - Formal analysis
      - Software
      - Methodology
      - Project administration
      - Validation
      - Writing - original draft
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Martin Cihak
    corresponding: false
    orcid: "0000-0001-7765-4799"
    email: "martin.cih@gmail.com"
    role:
      - Conceptualization
      - Investigation
      - Methodology
      - Writing - original draft
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Tereza Uhrova
    corresponding: false
    email: "tereza.uhrova@vfn.cz"
    role:
      - Investigation
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Robert Jech
    corresponding: false
    orcid: "0000-0002-9732-8947"
    email: "jech@cesnet.cz"
    role:
      - Funding acquisition
      - Resources
      - Writing - review & editing
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Ondrej Bezdicek
    corresponding: true
    orcid: "0000-0002-5108-0181"
    email: "ondrej.bezdicek@gmail.com"
    role:
      - Investigation
      - Data curation
      - Funding acquisition
      - Conceptualization
      - Project administration
      - Supervision
      - Writing - original draft
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
abstract: ""
keywords:
  - Parkinson’s disease
  - Parkinson's disease dementia
  - level I criteria
format:
 apaquarto-doc:
   documentmode: man
   a4paper: true
   pdf-engine: lualatex
   floatsintext: false
   keep-tex: false
bibliography: references.bib
---

{{< include _extensions/wjschne/apaquarto/_apa_title.qmd >}}

# Introduction

Parkinson’s disease (PD) is a neurodegenerative disorder characterized by a gradual and progressive onset of motor symptoms, including rigidity, bradykinesia, and resting tremors, which eventually extend to both motor and non-motor impairments [@postuma2015]. Beyond these hallmark features, cognitive decline is a critical aspect of the disease trajectory, culminating in Parkinson’s disease dementia (PDD) in a substantial subset of patients.

The diagnostic criteria for PDD, first formalized in 2006, were heavily influenced by frameworks established for Alzheimer’s disease (AD) due to the absence of PD-specific biomarkers that could support a biologically informed classification system [@emre2007; @dubois2007]. While effective at the time, these criteria lacked the granularity needed to address the unique pathophysiological and cognitive features of PD [@yamashita2023].

Subsequent advancements in the field have introduced the concept of Parkinson’s disease mild cognitive impairment (PD-MCI), providing a nuanced understanding of cognitive deficits in PD. The PD-MCI criteria, published in 2012, have propelled progress in diverse areas, including clinical characterization, genetic correlates, therapeutic strategies, clinical trial design, and the evaluation of progression rates to PDD [@litvan2012; @hoogland2017; @hoogland2019; @aarsland2021].

Currently, efforts are underway to refine the PDD diagnostic framework to enhance its consistency and utility in research and clinical contexts across multiple centers [@kulisevsky2024]. Our previous investigations have emphasized the importance of rigorous psychometric methodologies in differentiating PD-MCI from PDD, especially given that PDD remains a contraindication for deep brain stimulation (DBS) [@bezdicek2016; @deuschl2006].

The present study evaluates the diagnostic concordance between the original Level I PDD criteria, as outlined by the Movement Disorder Society (MDS) Task Force [@emre2007; @dubois2007] and the recently revised criteria [@kulisevsky2024] within a PD cohort selected for DBS. Specifically, we aimed to:

Compare diagnostic outcomes using the original and revised Level I criteria and assess the agreement of revised Level I criteria with comprehensive neuropsychological assessment at Level II.Compare diagnostic outcomes using the original and revised Level I criteria.

By addressing these objectives, the study seeks to validate the revised PDD criteria and explore their applicability in the context of DBS eligibility, thereby contributing to the refinement of cognitive assessment protocols in PD.

# Methods

## Participants

The study included retrospective data from patients diagnosed with idiopathic PD by a movement disorder specialist according to the MDS Clinical Diagnostic Criteria for PD [@postuma2015] Clinical records from January 2014 to December 2023 were analyzed. All participants underwent neuropsychological evaluations conducted by a trained clinical psychologist as part of routine cognitive assessments for determining eligibility for Deep Brain Stimulation (DBS) at the General University Hospital in Prague.

The study protocol was approved by the Ethics Committee of the General University Hospital in Prague. Informed consent was obtained from all patients prior to their neuropsychological assessments, in compliance with ethical research standards.

## Neuropsychological Assessment

Cognitive performance was evaluated using both the Mini-Mental State Examination (MMSE) [@stepankova2014; @folstein1975] and the Montreal Cognitive Assessment (MoCA) [@kopecek2016; @nasreddine2005] providing measures of global cognitive functioning. A comprehensive neuropsychological battery was administered in accordance with Level 2 criteria for Parkinson’s disease mild cognitive impairment (PD-MCI) proposed by the MDS Task Force [@litvan2012]. The details of this battery, including the regression-based calculator for normative scoring, have been outlined in a previous publication [@bezdicek2017].

In addition to other neuropsychological measures, the assessment incorporated the Clock Drawing Test (CDT) and Letter Fluency tasks to evaluate specific cognitive domains. For determining cognitive performance aligned with Level 1 criteria for Parkinson’s disease dementia (PDD), test scores defined by Dubois et al. [@dubois2007] nd their MoCA equivalents were utilized. Details of the scoring methods and their alignment with diagnostic criteria are summarized in Table A.
```{r}
#| label: tbl-martin
#| tbl-cap: Comparison of criteia.
#| apa-note: The note below the table.

tibble::tribble(
  ~"Criteria", ~"Former", ~"New",
  "", "MMSE < 26", "MoCA < 26", 
  "Attention", "Serial 7's/months reversed", "Serial 7's",
  "Executive functions", "Lexical Fluency (S)/ Clock Drawing Test", "Lexical Fluency (K)/ Clock Drawing Test",
  "Visuo-spatial functions", "Drawing of Pentagons", "Drawing of Cube",
  "Memory", "3-Word Recall", "5-Word Recall",
  "Language", "-", "Naming of Animals",
  ) |>
  flextable::flextable() |>
  flextable::theme_apa()
```

To measure independence in activities of daily living, we administered the Functional Activities Questionnaire (FAQ) [@bezdicek2011; @pfeffer1982]. To assess neuropsychiatric functioning, we used Beck Depression Inventory-II (BDI) [@ciharova2020; @beck1996] and State-Trait Anxiety Inventory (STAI) [@spielberg1983; @mullner1980]. Psychotic symptoms were assessed in an interview by a trained psychiatrist.

# Results

# Discussion

# References