### 2025-03-03 texts/theart.qmd
Rephrasing of the operationalization

---
lang: en
title: "Updated Criteria for the Diagnostic Procedure for Parkinson’s Disease Dementia on Level I and their Validity in Deep Brain Stimulation Cohort"
shorttitle: "Parkinson's Disease Dementia Level I Criteria"
author:
  - name: Martina Kvapilova
    corresponding: false
    orcid: "0009-0007-4665-3946"
    role:
      - Conceptualization
      - Data curation
      - Writing - original draft
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Josef Mana
    corresponding: false
    orcid: "0000-0002-7817-3978" 
    email: "josef.mana@protonmail.com"
    role:
      - Conceptualization
      - Data curation
      - Investigation
      - Formal analysis
      - Software
      - Methodology
      - Project administration
      - Validation
      - Writing - original draft
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Tereza Uhrova
    corresponding: false
    email: "tereza.uhrova@vfn.cz"
    role:
      - Investigation
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Robert Jech
    corresponding: false
    orcid: "0000-0002-9732-8947"
    email: "jech@cesnet.cz"
    role:
      - Funding acquisition
      - Resources
      - Writing - review & editing
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
  - name: Ondrej Bezdicek
    corresponding: true
    orcid: "0000-0002-5108-0181"
    email: "ondrej.bezdicek@gmail.com"
    role:
      - Investigation
      - Data curation
      - Funding acquisition
      - Conceptualization
      - Project administration
      - Supervision
      - Writing - original draft
    affiliations:
      - id: "1FM"
        name: "First Faculty of Medicine and General University Hospital in Prague, Charles University, Czech Republic"
        department: "Department of Neurology and Centre of Clinical Neuroscience"
abstract: ""
keywords:
  - Parkinson’s disease
  - Parkinson's disease dementia
  - level I criteria
format:
 apaquarto-docx:
   documentmode: man
   a4paper: true
   pdf-engine: lualatex
   floatsintext: false
   keep-tex: false
bibliography: references.bib
---

# Introduction

Parkinson’s disease (PD) is a neurodegenerative disorder characterized by a gradual and progressive onset of motor symptoms, including rigidity, bradykinesia, and resting tremors, which eventually extend to both motor and non-motor impairments [@postuma2015]. Beyond these hallmark features, cognitive decline is a critical aspect of the disease trajectory, culminating in Parkinson’s disease dementia (PDD) in a substantial subset of patients.

The diagnostic criteria for PDD, first formalized in 2006, were heavily influenced by frameworks established for Alzheimer’s disease (AD) owing to the absence of PD-specific biomarkers that could facilitate a biologically grounded diagnostic system [@emre2007; @dubois2007]. Although these criteria provided a valuable initial foundation, they lacked the specificity required to capture the distinct pathophysiological and cognitive features of PD-related dementia [@yamashita2023].

An important feature of the original criteria was the provision of an algorithm that allowed for flexibility in test selection [@dubois2007]. Specifically, clinicians could choose between months reversed or seven backwards for attention assessment, lexical fluency or clock drawing for executive function evaluation, MMSE pentagons for visuospatial ability, and three-word recall for memory assessment. The availability of multiple operationalization options for PDD enhances the ability to examine the psychometric properties of the construct. Agreement across different criteria allows for the parallel computation of inter-rater reliability , which, in turn, facilitates the calculation of construct validity [@conway1995] further strengthening the diagnostic framework for PDD.

Subsequent advancements have introduced the concept of Parkinson’s disease mild cognitive impairment (PD-MCI), refining the understanding of cognitive dysfunction in PD. The PD-MCI criteria, published in 2012, have propelled progress in diverse areas, enabling improvements in clinical characterization, identification of genetic correlates, therapeutic interventions, clinical trial design, and the assessment of progression risk to PDD [@litvan2012; @hoogland2017; @hoogland2019; @aarsland2021].

Currently, efforts are focused on refining the PDD diagnostic framework to improve its consistency and applicability in both research and clinical contexts across multiple centers [@kulisevsky2024]. Our prior investigations have underscored the importance of employing rigorous psychometric methodologies to differentiate PD-MCI from PDD, particularly given that a diagnosis of PDD remains a contraindication for deep brain stimulation (DBS) [@bezdicek2016; @deuschl2006].

The present study evaluates the diagnostic concordance between the original Level I PDD criteria, as established by the Movement Disorder Society (MDS) Task Force [@emre2007; @dubois2007] and the recently revised criteria [@kulisevsky2024] within a PD cohort selected for DBS. Specifically, we aimed to:

Compare diagnostic outcomes utilizing the original and revised Level I criteria and assess the agreement of revised Level I criteria with comprehensive neuropsychological assessment at Level II.Compare diagnostic outcomes using the original and revised Level I criteria.

By addressing these objectives, this study seeks to validate the revised PDD criteria and evaluate their relevance in the context of DBS eligibility, thereby contributing to the refinement of cognitive assessment protocols in PD.

# Methods

## Participants

This study retrospectively analyzed clinical data from patients diagnosed with idiopathic PD by a movement disorder specialist according to the MDS Clinical Diagnostic Criteria for PD [@postuma2015] Clinical records spanning January 2014 to December 2023 were examined. All participants underwent neuropsychological evaluations conducted by a trained clinical psychologist as part of standard preoperative cognitive assessments for DBS eligibility at the General University Hospital in Prague.

Ethical approval for the study protocol was obtained from the Ethics Committee of the General University Hospital in Prague. Informed consent was secured from all patients prior to their neuropsychological assessments, in adherence to ethical research guidelines.

## Neuropsychological Assessment

Cognitive performance was evaluated using both the Mini-Mental State Examination (MMSE) [@stepankova2014; @folstein1975] and the Montreal Cognitive Assessment (MoCA) [@kopecek2016; @nasreddine2005] providing measures of global cognitive functioning. A comprehensive neuropsychological battery was administered in accordance with the Level II criteria for PD-MCI, as delineated by the MDS Task Force  [@litvan2012]. The methodological details, including a regression-based normative scoring approach, have been described in previous study [@bezdicek2017].

In addition to core cognitive assessments, specific tasks such as the Clock Drawing Test (CDT) and Letter Fluency tasks were incorporated to evaluate domain-specific impairments. The diagnostic classification of PDD using Level I criteria was based on established scoring thresholds from Dubois et al. [@dubois2007] and their corresponding MoCA equivalents. A summary of the scoring methodologies and their alignment with diagnostic criteria is provided in Table A.

```{r}
#| label: tbl-martin
#| tbl-cap: Comparison of criteia.
#| apa-note: The note below the table.

tibble::tribble(
  ~"Criteria", ~"Former", ~"New",
  "", "MMSE < 26", "MoCA < 26", 
  "Attention", "Serial 7's/months reversed", "Serial 7's",
  "Executive functions", "Lexical Fluency (S)/ Clock Drawing Test", "Lexical Fluency (K)/ Clock Drawing Test",
  "Visuo-spatial functions", "Drawing of Pentagons", "Drawing of Cube",
  "Memory", "3-Word Recall", "5-Word Recall",
  "Language", "-", "Naming of Animals",
  ) |>
  flextable::flextable() |>
  flextable::theme_apa()
```

To assess functional impairment, the Functional Activities Questionnaire (FAQ) [@bezdicek2011; @pfeffer1982]. was administered. Additionally, neuropsychiatric status was evaluated using the Beck Depression Inventory-II (BDI) [@ciharova2020; @beck1996] and State-Trait Anxiety Inventory (STAI) [@spielberg1983; @mullner1980]. Psychotic symptoms were assessed through structured psychiatric interviews conducted by a trained psychiatrist.

## Operationalization of dementia

In this study, we applied three different types of criteria for diagnosing PDD on level I. The first was based on the original framework [@dubois2007], which was based on MMSE and included assessments of attention, executive function, visuospatial abilities, and memory. The second was based on the recent call for change of dementia diagnostic guidelines [@kulisevsky2024], refining cognitive domain assessments to improve sensitivity to PD-related dementia, with specific items from the MoCA utilized for this purpose. The third approach applied the Czech version of the shortened Montreal Cognitive Assessment (sMoCA) [@bezdicek2020], testing whether equivalent cognitive impairments can be reliably identified using a more time-efficient testing protocol.

To facilitate a clear comparison of these diagnostic frameworks, a structured table summarizing the components and applicability of each criterion is provided below.

Insert Table 2

# Results

# Discussion

# References